---
title: "PML-Course Project"
author: "Adrian CG"
date: "22 Feb 2015"
output:
  html_document:
    keep_md: yes
    theme: journal
bibliography: bibliography.bib
---

```{r downloadData, echo=FALSE, results='hide', cache=FALSE}
library("knitcitations"); cleanbib(); bib <- read.bibtex("bibliography.bib");
source("utils.R")
pmlTraining <- read.csv("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv")
pmlTesting <- read.csv("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv")
```

## Data Preprocessing
On the first glance into the data, we observe a lot of wrongly concluded data types for the scalar measurement columns (ranging from 8 to 159). Thus, we coerce them to the appropriate numeric classes:

```{r correctClasses, cache=TRUE}
for (i in 8:159) {
    pmlTraining[,i] <- as.numeric(pmlTraining[,i])
    pmlTesting[,i] <- as.numeric(pmlTesting[,i])
}
str(pmlTraining[,c(1:8,159,160)])
```

After manual inspection of the entries in the training set, we found some samples at the end of each execrise window containing a full set of the derived variables, while others (beginning and within windows) only contain the base measurements. For classification purposes of a whole movement as mentioned in the original paper `r citep(bib[["Velloso2013"]])`, the derived variables seem to be a better fit. However, sine they are rather sparse and can't be recomputed for single points within the exercise repetion, we constrain ourselves to the 93 variables that don't contain `NA`s.

```{r restrictMeasured, echo=TRUE, cache=TRUE}
useableColumns <- apply(!is.na(pmlTraining), 2, any)
pmlTraining <- pmlTraining[,useableColumns]
pmlTesting <- pmlTesting[,useableColumns]
```

## Prediction
### Data-Splitting

### Model Building

### Assessing Prediction Accuracy

### Application on Original Test-Dataset

## References
```{r bibiliography, echo=FALSE, cache=TRUE, message=FALSE, results='asis'}
print(bibliography(), "html")
```

## Appendix
### Raw Usable Parameter List
```{r usableParameters, echo=FALSE, cache=TRUE}
str(names(pmlTraining),vec.len=160)
```

